COVID-19 dataset strength

### 2020-07-16 covid_dataset_strength/covid_dataset_strength.Rmd
Taking unique sequences into account, the dataset is weak

Abbr  |Full
------|-----------
AA    |Amino acid
AAs   |Amino acids
Abbr  |Abbreviation
TMH   |Transmembrane helix
TMHs  |Transmembrane helices

@@

```{r cache=TRUE}
topology <- pureseqtmr::predict_topology(fasta_filename)
n_proteins <- nrow(topology)
n_tmhs <- sum(tally_tmhs(topology)$n_tmhs)
n_non_tmh_aas <- sum(stringr::str_count(topology$topology, "0"))
n_tmh_aas <- sum(stringr::str_count(topology$topology, "1"))
n_aas <- n_non_tmh_aas + n_tmh_aas
f_aas <- n_tmh_aas / (n_tmh_aas + n_non_tmh_aas)
proteome_size <- n_aas
```

@@

There are `r n_sequences` COVID-19 sequences (`https://www.gisaid.org/`).

```{r}
n_tmh_aa_mutations <- tmh_aa_evolution_rate_per_sequence * n_sequences
```

>>>>

Abbr  |Full
------|-----------
AA    |Amino acid
AAs   |Amino acids
Abbr  |Abbreviation
Abbrs |Abbreviations
TMH   |Transmembrane helix
TMHs  |Transmembrane helices

@@

```{r cache=TRUE}
topology <- pureseqtmr::predict_topology(fasta_filename)
```

Plot it:

```{r}
pureseqtmr::plot_topology(topology)
```

Extract data from the topology:

```{r}
n_proteins <- nrow(topology)
n_tmhs <- sum(tally_tmhs(topology)$n_tmhs)
n_non_tmh_aas <- sum(stringr::str_count(topology$topology, "0"))
n_tmh_aas <- sum(stringr::str_count(topology$topology, "1"))
n_aas <- n_non_tmh_aas + n_tmh_aas
f_aas <- n_tmh_aas / (n_tmh_aas + n_non_tmh_aas)
proteome_size <- n_aas
```

@@

There are `r n_sequences` COVID-19 sequences (`https://www.gisaid.org/`).
Not all of these are unique.
We will determine the number of unique sequences
from the percentage of unique sequence per protein.

To do so, we read all COVID-19 sequences:

```{r cache=TRUE}
all_sequences_filename <- "../downloads/allprot0621.fasta"
testthat::expect_true(file.exists(all_sequences_filename))
sequences <- seqinr::read.fasta(
  all_sequences_filename, 
  seqtype = "AA",
  as.string = TRUE
)
```

Grouping them by protein:

```{r}
t <- tibble::tibble(
  prot_name = as.character(
    na.omit(
      stringr::str_match(
        names(sequences),
        "^([:upper:][:alnum:]{0,5})\\|"
      )[, 2]
    )
  ),
  sequence = as.character(sequences)
)
library(dplyr)
t$prot_name <- as.factor(t$prot_name)
t_unique <- t %>% group_by(prot_name) %>% 
  summarise(n = n(), n_distinct = n_distinct(sequence)) %>%
  mutate(perc = n_distinct / n)
perc_unique <- mean(t_unique$perc)
```

Of all proteins, on average, `r perc_unique` are unique.

```{r}
n_unique_sequences <- perc_unique * n_sequences
```

Therefor, there are `r n_unique_sequences` unique sequences.

```{r}
n_tmh_aa_mutations <- tmh_aa_evolution_rate_per_sequence * n_unique_sequences
```
